Post-Disturbance Tree Community Trajectories in a Neotropical Forest

### 2018-04-19 WholePlotTrajectories.Rmd
Intro

In the short-term, diversity dynamics demonstrated negligible or even positive impacts of disturbance on communities diversity, and have been formalized by the intermediate disturbance hypothesis (IDH) that predicts a maximized species diversity at intermediate level of disturbance[@Molino2001;@Kariuki2006a;@Berry2008a]. 
The IDH however has scarcely been tested in the long term and mainly relied on the analysis of species richness that gives limited or misleading information on forests recovery and functioning [@Martin2015;@Chaudhary2016].
Communities composition and evenness are crucial to resolve conservation issues and assess the ecological rules shaping communities abundance distribution [@Magurran1988;@Lavorel2002;@Bellwood2006].
A promosing perspective is also the functional approach accounting for species biological attributes related to individuals fitness and ecosystem properties [@Violle2007b;@Moretti2009;@Baraloto2012a;@Scheiter2013].
Major functional traits revealing species ecology and performance were largely adopted through a workable framework assessing forests dynamics and functioning [@Diaz2005;@Villeger2008a].
This approach for example highlighted in tropical rainforests the environmental filters fostering disturbance resistant species with rapid growth and efficient resources acquisition [@Molino2001;@Haddad2008].
This was translated by the shift from "conservative" species with long-lived, dense tissues growing slowly but with scarce resources, to "acquisitive" fast-growing species with light tissues and efficient resources use [@TerSteege2001;@Reich2014;@Herault2011].
Such shift would be detected with a combination of resource acquisition funcitonal traits (leaf area, density and chlorophyll content and stem specific gravity and bark thickness) and growth and reproduction life history traits (seed mass and maximum height) already proved relevant [@Wright2004;@Westoby2006a;@Chave2009b].
A functional approach combined to the analysis of taxonomic diversity and composition is then a reliable way to validate the IDH in the long term and from communities evenness and functinal perspective.
A full interpretation of communities response would be completed with communities taxonomic and functional composition that would elucidate the determinism of forest recovery and the maintenance of intrinsinc differences. 
Those development would be as much information to understand the mechanisms of forests response to diversity and help conservation and management strategies.

Here we investigated over 30 years the response of 75 ha of forests plots set up on a gradient of disturbance intensity, from 10 to 60% of ecosystem biomass removed.
We made use of a large functional traits database browsing major leaf, stem and seed traits and species maximum height to draw communities taxonomic and functional composition and diversity trajectories after disturbance.
Specifically, we (i) tested the validity of the IDH arguing for enhanced diversity after moderate disturbance and (ii) confirmed the changes in ecological rules shaping communities, from competitive exclusion between "conservative", slow-growing tree species to environmental filtering of "acquisitive", fasting-growing ones.
Eventually we examined the different trajectories to (iii) resolve the resilience of tropical forests and question their future dynamics regarding their state 30 years after disturbance.

>>>>

In the short-term diversity dynamics demonstrated negligible or even positive impacts of disturbance on communities diversity, which have been formalized by the intermediate disturbance hypothesis (IDH) stating a maximized species diversity at intermediate disturbance intensity [@Molino2001;@Kariuki2006a;@Berry2008a]. 
However, validations of the IDH in the long term remain scarce and mainly rely on the analysis of species richness that gives limited or misleading information on forests recovery and functioning [@Martin2015;@Chaudhary2016].
More releveant monitoring would encompass communities composition, that are crucial for conservation issues, and evenness, that revealing ecological rules underlying communities' structure [@Magurran1988;@Lavorel2002;@Bellwood2006].
Furthermore, the functional approach accounting for species biological attributes would be insightful as it reveals species fitness and ecosystems functioning [@Violle2007b;@Moretti2009;@Baraloto2012a;@Scheiter2013].
In that respect major functional traits related to species ecology and performance were largely adopted through a relevant framework [@Diaz2005;@Villeger2008a].
The functonial trait-based approach, for example, highlighted in tropical rainforests the environmental filters fostering disturbance resistant species with rapid growth and efficient resources acquisition [@Molino2001;@Haddad2008].
This translated by shifts from "conservative" slow-growing species dealing with scarce resources, to "acquisitive" fast-growing species with rapid and efficient use of abundant resources [@TerSteege2001;@Reich2014;@Herault2011].
It was mirrored by shifts in key functional traits related to resource acquisition (leaf area, density and chlorophyll content, and stem specific gravity and bark thickness), tree growth and reproduction life history traits (seed mass and maximum height) [@Wright2004;@Westoby2006a;@Chave2009b].
A proper monitoring of communities response should therefore encompass taxonomic and functional diversity and composition measures to test the validity of the IDH in the long term for tropical forests, and clarify the resilience of communities evenness, composition,and also functioning.
The trajectories followed by all these facets would highlight the role of deterministic processes, like competitive exclusion or abiotic selection, and the communities' convergence maintaining intrinsinc differences in diversity and composition,which is as much insgights for future adaptive conservation strategies [@Adler2007].

Here we investigated over 30 years the response of 75 ha of forests plots set up on a gradient of disturbance intensity, from 10 to 60% of ecosystem biomass removed.
We made use of a large functional traits database browsing major leaf, stem and seed traits and species maximum height to draw the trajectories over time of communities taxonomic and functional composition and diversity.
Specifically, we (i) tested the validity of the IDH in the long term for tropical hiperdiverse forest and highlighted the ecological rules shaping their response to disturbance, (ii) clarified the different facets of communities resilience in terms of communities composition, diversity and functioning (iii) questioned the completeness of communities recovery given the altered functional redundancy.
>>revoir la fin

### 2018-04-19 WholePlotTrajectories.tex
Intro

In the short-term, diversity dynamics demonstrated negligible or even
positive impacts of disturbance on communities diversity, and have been
formalized by the intermediate disturbance hypothesis (IDH) that
predicts a maximized species diversity at intermediate level of
disturbance\citep{Molino2001, Kariuki2006a, Berry2008a}. The IDH however
has scarcely been tested in the long term and mainly relied on the
analysis of species richness that gives limited or misleading
information on forests recovery and functioning
\citep{Martin2015, Chaudhary2016}. Communities composition and evenness
are crucial to resolve conservation issues and assess the ecological
rules shaping communities abundance distribution
\citep{Magurran1988, Lavorel2002, Bellwood2006}. A promosing perspective
is also the functional approach accounting for species biological
attributes related to individuals fitness and ecosystem properties
\citep{Violle2007b, Moretti2009, Baraloto2012a, Scheiter2013}. Major
functional traits revealing species ecology and performance were largely
adopted through a workable framework assessing forests dynamics and
functioning \citep{Diaz2005, Villeger2008a}. This approach for example
highlighted in tropical rainforests the environmental filters fostering
disturbance resistant species with rapid growth and efficient resources
acquisition \citep{Molino2001, Haddad2008}. This was translated by the
shift from ``conservative'' species with long-lived, dense tissues
growing slowly but with scarce resources, to ``acquisitive''
fast-growing species with light tissues and efficient resources use
\citep{TerSteege2001, Reich2014, Herault2011}. Such shift would be
detected with a combination of resource acquisition funcitonal traits
(leaf area, density and chlorophyll content and stem specific gravity
and bark thickness) and growth and reproduction life history traits
(seed mass and maximum height) already proved relevant
\citep{Wright2004, Westoby2006a, Chave2009b}. A functional approach
combined to the analysis of taxonomic diversity and composition is then
a reliable way to validate the IDH in the long term and from communities
evenness and functinal perspective. A full interpretation of communities
response would be completed with communities taxonomic and functional
composition that would elucidate the determinism of forest recovery and
the maintenance of intrinsinc differences. Those development would be as
much information to understand the mechanisms of forests response to
diversity and help conservation and management strategies.

Here we investigated over 30 years the response of 75 ha of forests
plots set up on a gradient of disturbance intensity, from 10 to 60\% of
ecosystem biomass removed. We made use of a large functional traits
database browsing major leaf, stem and seed traits and species maximum
height to draw communities taxonomic and functional composition and
diversity trajectories after disturbance. Specifically, we (i) tested
the validity of the IDH arguing for enhanced diversity after moderate
disturbance and (ii) confirmed the changes in ecological rules shaping
communities, from competitive exclusion between ``conservative'',
slow-growing tree species to environmental filtering of ``acquisitive'',
fasting-growing ones. Eventually we examined the different trajectories
to (iii) resolve the resilience of tropical forests and question their
future dynamics regarding their state 30 years after disturbance.

>>>>

In the short-term diversity dynamics demonstrated negligible or even
positive impacts of disturbance on communities diversity, which have
been formalized by the intermediate disturbance hypothesis (IDH) stating
a maximized species diversity at intermediate disturbance intensity
\citep{Molino2001, Kariuki2006a, Berry2008a}. However, validations of
the IDH in the long term remain scarce and mainly rely on the analysis
of species richness that gives limited or misleading information on
forests recovery and functioning \citep{Martin2015, Chaudhary2016}. More
releveant monitoring would encompass communities composition, that are
crucial for conservation issues, and evenness, that revealing ecological
rules underlying communities' structure
\citep{Magurran1988, Lavorel2002, Bellwood2006}. Furthermore, the
functional approach accounting for species biological attributes would
be insightful as it reveals species fitness and ecosystems functioning
\citep{Violle2007b, Moretti2009, Baraloto2012a, Scheiter2013}. In that
respect major functional traits related to species ecology and
performance were largely adopted through a relevant framework
\citep{Diaz2005, Villeger2008a}. The functonial trait-based approach,
for example, highlighted in tropical rainforests the environmental
filters fostering disturbance resistant species with rapid growth and
efficient resources acquisition \citep{Molino2001, Haddad2008}. This
translated by shifts from ``conservative'' slow-growing species dealing
with scarce resources, to ``acquisitive'' fast-growing species with
rapid and efficient use of abundant resources
\citep{TerSteege2001, Reich2014, Herault2011}. It was mirrored by shifts
in key functional traits related to resource acquisition (leaf area,
density and chlorophyll content, and stem specific gravity and bark
thickness), tree growth and reproduction life history traits (seed mass
and maximum height) \citep{Wright2004, Westoby2006a, Chave2009b}. A
proper monitoring of communities response should therefore encompass
taxonomic and functional diversity and composition measures to test the
validity of the IDH in the long term for tropical forests, and clarify
the resilience of communities evenness, composition,and also
functioning. The trajectories followed by all these facets would
highlight the role of deterministic processes, like competitive
exclusion or abiotic selection, and the communities' convergence
maintaining intrinsinc differences in diversity and composition,which is
as much insgights for future adaptive conservation strategies
\citep{Adler2007}.

Here we investigated over 30 years the response of 75 ha of forests
plots set up on a gradient of disturbance intensity, from 10 to 60\% of
ecosystem biomass removed. We made use of a large functional traits
database browsing major leaf, stem and seed traits and species maximum
height to draw the trajectories over time of communities taxonomic and
functional composition and diversity. Specifically, we (i) tested the
validity of the IDH in the long term for tropical hiperdiverse forest
and highlighted the ecological rules shaping their response to
disturbance, (ii) clarified the different facets of communities
resilience in terms of communities composition, diversity and
functioning (iii) questioned the completeness of communities recovery
given the altered functional redundancy.
\textgreater{}\textgreater{}revoir la fin
